---
title: "data-viz"
author: "Wylie Hampson"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General Use
library(tidyverse)
library(here)
library(janitor)

# Specifically for dataviz:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)
```

## Get the liozards data

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

Histogram (really bad one) of lizard weights

```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "orange",
                 color = "purple",
                 size = 1, 
                 linetype = "dotdash")
```

```{r}
# Total length on x-axis and wight on y-axis
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             fill = "purple",
             color = "black",
             size = 5,
             alpha = 0.1)
```

Bubble plot where the color of the points changes based on common_name and the size of points varies based on total_length.

DISCLAIMER: BAD IDEA

```{r}
ggplot(data = lizards,  aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length,
                 fill = "black"),
             shape = 22) +
  theme_dark()
```

```{r}
ggplot(data = lizards,  aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name, ncol = 5)
```

```{r}
ggplot(data = lizards,  aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)
```

Find total lizard counts by common_name

```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

# Does the same thing
lizard_counts <- lizards %>% dplyr::count(common_name)

lizard_counts_cn_tail <- lizards %>% count(common_name, tail)
```

```{r}
ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, n), x = n)) +
  geom_col(aes(fill = common_name)) +
  labs(x = "Lizard Count",
       y = "Common Name")
```

Try converting common_name to an ordered factor outside of ggplot

```{r}
lizard_counts <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, n))
```

## Axis scale breaks, limits, and labels

Scatterplot: total_length (x) vs. weight (y)

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0 , 10, 50, 500),
                     limits = c(0, 500),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

Transform the date column to class date, then find counts of observations by date

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)

class(lizard_counts$date)
```

Make a line plot with geom_line of date (x) and count (y)

```{r}
ggplot(data = lizard_counts, aes(x = date, y = n)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%y")
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_stepsn(colors = c("red", "yellow", "green", "pink"), 
                     breaks = c(0, 20, 40, 50))
```

### Update a color scheme using a palette in paletteer
Make a horizontal box plot with common_name on the y-axis, total_length on the x-axis, with color changing based on common_name

```{r}
lizards_fct <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

ggplot(data = lizards_fct, aes(x = total_length, y = common_name)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "vapoRwave::mallSoft")
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    panel.grid = element_blank(),
    plot.background = element_rect(fill = "cyan4"),
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "yellow",
                              size = 15)
  )
```

## ggrepel

Make a subset from lizards, called ww_lizards, that only contains observations for "western whiptail" lizards from the site "sand"

```{r}
ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(data = ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num), size = 3, max.overlaps = 20)
  
```

Create a scatterplot of gdpPerCap (x) and lifeExp (y), labeled by country name, filter for only europe.

```{r}
gapminder_practice <- gapminder %>% 
  filter(continent == "Europe", year == 2002)

ggplot(gapminder_practice, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)
```

## gghighlight

```{r}
p <- ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)
```

```{r}
q <- ggplot(lizards, aes( x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) +
  gghighlight(max(weight) > 30)

q
```

## patchwork for compound figures

```{r}
(p | q) / q & 
  theme_minimal()
```

## A few new graph types

### Marginal Plots

```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

ggplot(whiptails, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_rug()
```

marginal plot with boxplots in the margins

```{r}
my_plot <- ggplot(whiptails, aes(x = total_length, y = weight)) +
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = c("cyan4", "black", "goldenrod"), 
                     name = "Sex:",
                     labels = c("Female", "Juvenile", "Male")) +
  theme_minimal() +
  theme(
    legend.position = "bottom"
  )
  
ggMarginal(my_plot, type = "boxplot", groupColour = TRUE)
```

### Beeswarm plot

```{r}
ggplot(data = whiptails, aes(x = sex, y = weight)) +
  geom_beeswarm() +
  geom_boxplot(fill = NA) +
  geom_violin(fill = NA)
```

### A heatmap with geom_title()

Make a new dataframe called lizard_counts, starting with lizards, with date converted to class(date), then count by year and common_name the number of lizards observed.

```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  mutate(year = lubridate::year(date)) %>% 
  group_by(year, common_name) %>% 
  summarize(count = n())
```

```{r}
ggplot(lizard_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = count)) +
  geom_text(aes(label = count), color = "white", size = 3) +
  scale_fill_viridis_c()
```

### Make a map!

Use read_sf to read in the doc.kml file

```{r}
jornada_vegetation <- read_sf(here::here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()

ggplot(data = jornada_vegetation) +
  geom_sf(aes(fill = name), color = NA) +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Dominant Vegetation:") +
  theme_minimal()
```






